---
title: "Initial Korean covid 19"
author: "Dorothy Decker"
date: "6/21/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?


## Methods

We will Investigate whether men and women are equally likely to die of COVID 19 from data collected in South Korea through May 2019. We will analyze the Sex of the patients and the result if they are alive or dead. The result is one of the variables. There is a factor variable $Sex and $sex with two levels, we have a helper file to use as a template to approach this problem. 




## Graphical results - use ggplot commands to do this  



#### Plot barplots ggplot way

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
Looking at the graphical results, we can assume there are more males dying than females. It looks like there are twice as many males, but in the first place there were more females than males. This virus is somehow impacting males more than females. 




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Using the numerical results, we can further interpret that males are dominating the section of death in comparison to females. In the first chart, there are 36% females dead, where there are 64% dead males. In the second chart, 1.4% are dead and there are 3% dead of males, so there isn't many people dying, but of this data set the gender of those dying are more male than female.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```



Using the inferential results, we can further conclude that although there is a smaller group of males than females, the males are still dying more than females.  The degrees of freedom is 1, while the result is 1. The P-value is 0.0012 is a small chance of showing there is no difference. With the Fisher's Exact Test the P-Value is 9e-04 the odds ratio is 2.3, which says the odds of a female living is 2.3 times greater than a male living in this specific study.  


## Conclusions


In conclusion, from examining all of the results, we can conclude that males are dying more than females. In this specific data set, COVID-19 has a greater impact on males. There were more females in the sample group, but regardless, the male group still had more deaths in comparison to the females. By looking at the graphical results, we can conclude that males are dying at a greater rate. As well as looking at the numerical and inferential results. Looking at the numbers and probability, just further proves that males are for some reason more impacted than females. 

